\section{The Twin Paradox}

\instructornote{%
By Matt Trawick.  Time: 30 minutes

}


\makelabheader %(Space for student name, etc., defined in master.tex)

\bigskip

\textit{The point of this exercise is to step you through a visualization of the twin paradox.  We see that a lot of the aging of Bob in Anna's frame happens when Anna turns around, when her frame is not inertial.}

Open the Mathematica file \filename{twin\_paradox.nb}, hit \button{Ctrl-A} to select all, and hit \button{Shift-Enter} to evaluate all the selected statements.  Pretend that the time axis has units in years.

Here's the story: Two twins (Bob and Anna) are separated at birth. Bob stays at home, and Anna flies away from the Earth at $0.6c$ until she reaches a star 30~light years away (as measured by Bob), then turns around and comes back at $0.6c$. The red and purple lines in the Minkowski diagram represent the worldlines of Bob and Anna, respectively.

\begin{enumerate}[wide]
\item  From the graph, how long does Anna's trip away from Earth (not counting the return) take as measured in Bob's frame? How long is the trip as measured in Anna's outgoing frame?
\answerspace{0.8in}

\item The blue lines in the graph are ``simultaneity lines'' in Bob's frame; everything along a single blue line happens simultaneously according to Bob or any observer in Bob's frame. According to Bob, when Bob has aged by 25 years, how much has Anna aged by? Answer by following the blue simultaneity line as you switch reference frames with the slider to the frame where Anna is at rest. (An answer by eyeballing the graph is fine.) \label{bob_aged_25}
\answerspace{0.8in}

\item The green lines in the graph are simultaneity lines in Anna's outgoing frame. According to Anna, when Anna has aged by 25 years, how much has Bob aged by?  Answer by following a green simultaneity line in the frame where Bob is at rest. \label{anna_aged_25}
\answerspace{0.8in}

\item Are your answers in parts~\ref{bob_aged_25} and \ref{anna_aged_25} the same, or different? Are the roles of Bob and Anna in this series of events symmetric or asymmetric, so far?
\answerspace{0.8in}

\pagebreak[3]
Now look at the second graph in the file. It shows the same events, but with the origin defined as the event of Anna turning around.  The twins are now born at $t=-50$~y (Bob's frame) and $t=-40$~y (Anna's frame).
\medskip

\item In Anna's outgoing reference frame, at the instant just \textit{before} Anna turns around, how old is Bob?  

\textit{Hint: You can follow a green simultaneity line to where it intersects Bob's worldline, marking an event in Bob's life that is simultaneous, in Anna's frame, with Anna being just about to turn around.  
Maybe Bob blows out some birthday candles that day.}

\textit{But ``how old is Bob, in Anna's reference frame'' still means Bob's biological age---how many candles he blew out, or, if he were a tree, how many annual rings he's grown.  That time is measured in Bob's frame.}

\answerspace{0.8in}

\item In the second graph, there is an additional cyan line. In which reference frame does this represent a simultaneity line?  (Possible choices are Bob's frame, Anna's outgoing frame, or Anna's return frame.)
\answerspace{0.8in}

\item In Anna's returning reference frame, at the instant just \textit{after} Anna turns around, how old is Bob?  
\answerspace{0.8in}

\item According to Anna, how much has Bob aged during the very short time it took her to turn around?
\answerspace{0.8in}

\item According to Bob, how much does Anna age during the time she turns around?
\answerspace{0.8in}

\textit{The non-inertial nature of Anna's frame means that Bob ages a lot during her turn-around, and this resolves the paradox.  When the whole thing is over and Anna returns to Earth, Bob really has aged more than Anna in both frames.  Bob's frame is inertial, and Anna's frame is not, due to her acceleration.  That's what makes their situations not symmetric.}


%Removed by Jack:
%\item From Bob's point of view, Anna moved away from him and then came back, so she should be younger.  But from Anna's point of view, Bob was the one moving away from her, and then later Bob was the one who was moving back towards her.  Should Anna expect Bob to be younger?  When the twins are reunited, they must both agree which twin looks younger.  What is it specifically that ``breaks the symmetry'' between their roles and determines who is actually younger? 
%I think neither my original nor Jack's revision really gets to the heart of the matter.  Let's rewrite this paragraph.
%\answerspace{1.1in}

\end{enumerate}
