\section{Skill Builder Problems in Relativity (More Fun with Anna and Bob!)}
\sectionmark{Skill Builder Problems} %shorter version of title for the header
%By Matt Trawick

%\makelabheader %(Space for student name, etc., defined in master.tex)

\SetEnumitemKey{SkillBuilderParts}{nosep,label=(\emph{\roman*})} %defines labparts to number parts as (a), (b), etc.

\bigskip

\textit{Students often struggle to solve complex problems involving multiple events and worldlines viewed in different reference frames.  Although there is no single recipe that can be applied to all relativistic kinematics problems, many problems involve specific skills that you can isolate and practice.  
In this set of exercises, you will be applying specific skills to the three scenarios below.
}
\bigskip
\begin{quote}

\textbf{Scenario A:}\\
At time $t=0$, Bob flies away from Earth in his spacecraft at $0.6 c$.  After one year, Anna, who has stayed on Earth, realizes that Bob has forgotten his physics textbook at home.  She grabs the book and chases after him at speed $0.9 c$.  Once she catches up to him and gives him the book, she immediately turns around and returns home at $0.8 c$.  \textit{All quantities given are in the reference frame of the Earth.}
\bigskip

\textbf{Scenario B:}\\
The starship Enterprise is returning to Earth at velocity $-0.4 c$.  At time $t=0$, when the Enterprise is 4~light years away from Earth, the ship sends a radio message towards Earth (at speed $c$) to announce that they are coming.  A supply vessel is already en route from the Earth to the Enterprise at $0.6 c$, and they intercept the radio message when the supply vessel is exactly 2~light years away from Earth.  The supply vessel continues towards the Enterprise at the same speed until the two ships rendezvous. \textit{All quantities given are in the reference frame of the Earth.}
\bigskip

\textbf{Scenario C:}\\
Anna buys a cheap used car from Craigslist and installs a hyperdrive engine in it, enabling the car to travel at $0.5 c$.  She arranges a drag race with Carl and his new roadster, which has a top speed of $0.75 c$.  Bob stands at the starting line and gives them the signal to go at $x=0$ and $t=0$.  (Unrealistic though it is, we'll imagine that both cars accelerate instantaneously and maintain their top speeds throughout the race.)

Unfortunately, Anna was too focused on her car's engine, and she neglected some other details.  As she races, Bob witnesses various parts fall off the car (\textit{All times given are in Bob's reference frame.}): 
\begin{itemize}[nosep]
\item The rear bumper falls off at $ct=100$ meters. 
\item The muffler falls off at $ct=200$ meters.
\item The hood flies off at $ct=400$ meters.
\end{itemize}  
At time $600~{\rm meters}/c$, Carl beams a text message at speed $c$ directly back to Anna to taunt her.  At the same time (in Bob's reference frame), Anna beams a text message at speed $c$ towards Bob, offering him a good deal on a late model Corolla.

\end{quote}

\vspace{0.5 in}

\textbf{Skill \#1: Draw Minkowski diagrams showing events and worldlines}

1A. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario A} in the Earth reference frame.  
\bigskip

1B. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario B} in the Earth reference frame.  
\bigskip

1C. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario C} in Bob's reference frame.  
\bigskip


\pagebreak


\textbf{Skill \#2: Calculate relationships between events using regular, smelly old intro physics equations like \boldmath$ \Delta t=t_{\rm final}-t_{\rm initial}$ and \boldmath${\Delta x = v \Delta t}$.}

2A. Answer these questions based on the events in \textbf{Scenario A}.
\begin{enumerate}[SkillBuilderParts]
\item In the Earth reference frame, how many light years does Bob travel in two years?
\answerspace{0.3in}

\item In the Earth reference frame, what is Anna's position at time $t=1.5$ years?
\answerspace{0.3in}

\item Write an equation giving Anna's position $x_A(t)$ in the Earth reference frame as a function of time after she leaves Earth.
\answerspace{0.3in}

\item At what time in the Earth frame does Anna catch Bob?  How far are they from Earth when they meet?
\answerspace{0.3in}

\item In the Earth frame, for how long is Anna away from Earth before she returns home?
\answerspace{0.4in}

\end{enumerate}

2B. Answer these questions based on the events in \textbf{Scenario B}.
\begin{enumerate}[SkillBuilderParts]
\item At what time in the Earth reference frame does the supply vessel intercept the Enterprise's message?
\answerspace{0.4in}

\item How far from the Earth (in the Earth reference frame) is the Enterprise when it meets the supply vessel?
\answerspace{0.4in}

\item How far (in the Earth reference frame) does the supply vessel travel between the time it intercepts the message and the time it meets the enterprise?
\answerspace{0.4in}

\end{enumerate}

2C. Answer these questions based on the events in \textbf{Scenario C}.
\begin{enumerate}[SkillBuilderParts]
\item How many meters from Bob (in Bob's reference frame) does the muffler fall off?
\answerspace{0.4in}

\item Where is Carl when the hood flies off?
\answerspace{0.4in}

\item In Bob's reference frame, where is Anna when she receives Carl's text message?
\answerspace{0.4in}
  
\end{enumerate}

\pagebreak

\textbf{Skill \#3: Visualize how Minkowski diagrams change under Lorentz transformation}

3A. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario A} in Bob's reference frame.  
\bigskip

3B. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario B} in the supply vessel's reference frame.  
\bigskip

3C. Draw a qualitative Minkowski diagram of the events in \textbf{Scenario C} in Anna's reference frame.  
\bigskip
\bigskip




\textbf{Skill \#4: Recognize proper times}

4A. Answer these questions based on the events in \textbf{Scenario A}.
\begin{enumerate}[SkillBuilderParts]
\item Who measures the proper time between Anna leaving Earth and Anna meeting Bob?
\answerspace{0.3in}

\item Who measures the proper time between Bob leaving home and Anna catching up to him? 
\answerspace{0.3in}

\item Who measures the proper time between Anna leaving Earth and Anna returning?  (Careful, this one's tricky!)
\answerspace{0.3in}

\end{enumerate}
4B. Answer these questions based on the events in \textbf{Scenario B}.
\begin{enumerate}[SkillBuilderParts]
\item Who measures the proper time between the Enterprise sending the radio message and the rendezvous between the Enterprise and the supply ship?
\answerspace{0.3in}

\item Who measures the proper time between the supply vessel intercepting the message and the supply vessel meeting the Enterprise?
\answerspace{0.3in}

\end{enumerate}

4C. Answer these questions based on the events in \textbf{Scenario C}.
\begin{enumerate}[SkillBuilderParts]
\item Who measures the proper time between Anna starting the race signal and Bob receiving Anna's text?  
\answerspace{0.3in}

\item Who measures the proper time between Anna's rear bumper falling off and Anna's muffler falling off?
\answerspace{0.3in}

\item During the race, who measures the proper length of Anna's car?
\answerspace{0.3in}

\end{enumerate}

\pagebreak

\textbf{Skill \#5: Calculate new coordinates for events, as well as distances and time lapses, under Lorentz transformations.}

\textit{In some cases, you can save a step by recognizing the proper length or time and multiplying or dividing by $\gamma$.}

5A. Answer these questions based on the events in \textbf{Scenario A}.
\begin{enumerate}[SkillBuilderParts]
\item In Bob's reference frame, at what time does Anna realize that Bob forgot his book?
\answerspace{0.3in}

\item In Bob's reference frame, at what time does Anna reach Bob?
\answerspace{0.3in}

\item In Bob's reference frame, how much time did it take Anna to reach him after she left Earth?
\answerspace{0.3in}

\item In Bob's frame, how far is Bob from Earth when Anna catches him?
\answerspace{0.3in}

\end{enumerate}

5B. Answer these questions based on the events in \textbf{Scenario B}.
\begin{enumerate}[SkillBuilderParts]
\item In the reference frame of the supply ship, at what time does the Enterprise send the radio message?
\answerspace{0.3in}

\item In the reference frame of the supply ship, how long does it take to reach the Enterprise after intercepting the message?
\answerspace{0.3in}

\item In the reference frame of the supply ship, how far does the Enterprise travel from the time it sends the radio message to the time it meets the supply ship?
\answerspace{0.3in}
\end{enumerate}

5C. Answer these questions based on the events in \textbf{Scenario C}.
\begin{enumerate}[SkillBuilderParts]
\item In Anna's reference frame, how much time passes between the muffler falling off and the hood flying off?
\answerspace{0.3in}

\item In Anna's reference frame, how far away from Bob is she when the hood flies off?
\answerspace{0.3in}

\item In Anna's reference frame, which happens first: Carl sending the text to Anna, or Anna sending the text to Bob?
\answerspace{0.3in}

\end{enumerate}

\pagebreak

\textbf{Skill \#6: Calculate absolute and relative velocities in new reference frames}

6A. Answer these questions based on the events in \textbf{Scenario A}.
\begin{enumerate}[SkillBuilderParts]
\item In the Earth's reference frame, what is Anna's velocity on the return trip relative to Bob?
\answerspace{0.6in}

\item In Bob's frame, what is Anna's velocity on her return trip?
\answerspace{0.6in}

\item In Bob's frame, what is Anna's velocity relative to the Earth on her return trip?
\answerspace{0.6in}

\end{enumerate}

6B. Answer these questions based on the events in \textbf{Scenario B}.
\begin{enumerate}[SkillBuilderParts]
\item In the reference frame of the Earth, what is the relative speed of the two ships as they approach each other?
\answerspace{0.6in}

\item In the reference frame of the Enterprise, what is the relative speed of the two ships as they approach each other ?
\answerspace{0.6in}

\item in the reference frame of the Enterprise, what is the speed of the supply ship relative to the Earth?
\answerspace{0.6in}
\end{enumerate}

6C. Answer these questions based on the events in \textbf{Scenario C}.
\begin{enumerate}[SkillBuilderParts]
\item In the reference frame of Anna, what is Bob's velocity?
\answerspace{0.6in}

\item In the reference frame of Anna, what is Carl's velocity?
\answerspace{0.6in}

\item In the reference frame of Bob, what is Carl's velocity?
\answerspace{0.6in}

\item In the reference frame of Anna, what is Carl's velocity relative to Bob?
\answerspace{0.6in}
\end{enumerate}


\pagebreak
\answerspace{0.6in}