\section{Proper Length, Proper Distance, and Spacetime Intervals}
%By Matt Trawick.

\makelabheader %(Space for student name, etc., defined in master.tex)

\bigskip

Open the file \filename{proper\_length\_vs\_proper\_distance.nb} in Mathematica. If you get a pop-up error message, you may need to click \button{Enable Dynamic Content}. Type \button{Ctrl-A} to select all lines, and hit \button{Shift-Enter} to execute them.

The graph you see represents a spacetime diagram (or ``Minkowski diagram'') of a train moving at high speed. The two purple lines represent the worldlines of the two ends of the train. The four colored dots represent specific events that occur at either the front end or the back end of the train.

All of the questions below can be answered by estimating values from the Minkowski diagram in various reference frames; there are no calculations required.

\textbf{Activity 1: Proper length and proper distance}

\begin{enumerate}[labparts]
\item Roughly what is the proper length of the train?
\answerspace{0.6in}

\item What is the proper distance between the red event and the green event?
\answerspace{0.6in}

\item What is the proper distance between the red event and the black event?
\answerspace{0.6in}

\item What is the proper distance between the red event and the light blue event?
\answerspace{0.6in}

\item Write a definition of proper length
\answerspace{0.7in}

\item Write a definition of proper distance
\answerspace{0.7in}

\item If two events occur at two ends of an object, is the proper distance between the events always the same as the proper length of the object itself?
\answerspace{0.7in}
\end{enumerate}

\textbf{Activity 2: Causality and Spacetime Intervals}
\begin{enumerate}[labparts]
\item Is the spacetime interval between the red event and the green event timelike, spacelike, or lightlike?  How do you know?
\answerspace{0.6in}

\item Is the spacetime interval  between the green event and the black event timelike, spacelike, or lightlike? 
\answerspace{0.6in}

\item Is the spacetime interval  between the red event and the light blue event timelike, spacelike, or lightlike?
\answerspace{0.6in}

\item Is it possible that the light blue event caused the red event?
\answerspace{0.6in}

\item Is it possible that the green event caused the black event?
\answerspace{0.6in}
\end{enumerate}

\textbf{Activity 3: When proper time and proper distance don't make sense}
\begin{enumerate}[labparts]
\item What is the proper time between the green event and the black event?
\answerspace{0.6in}

\item What is the proper time between the green event and the red event?  \textit{Note: trick question!}
\answerspace{0.6in}

\item What is the proper distance between the green event and the black event?
\answerspace{0.6in}

\item When can you not calculate the proper time between two events, and when can you not calculate the proper distance between two events?
\answerspace{0.6in}
\end{enumerate}
